Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2024-02-15 main_exciting.tex
Resubmission PRFluids

\noindent {\it Introduction}. Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{grossFluidFilmLubrication1980a,szeriFluidFilmLubrication2010,quereFluidCoatingFiber1999,dasilvasobrinhoStudyDefectsUltrathin1999,singhInkjetPrintingProcess2010,wijshoffDynamicsPiezoInkjet2010}. 
Moreover, the continuously growing interest for lab-on-a-chip devices~\cite{samieiReviewDigitalMicrofluidics2016,fockeLabonaFoilMicrofluidicsThin2010} as well as for printable electronics or printable photovoltaics~\cite{luechingerGraphenestabilizedCopperNanoparticles2008,ronsinFormationCrystallineBulk2022}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{buttAdaptiveWettingAdaptation2018}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{xinReversiblySwitchableWettability2010}, involving smart materials such as polymer brushes~\cite{stuartEmergingApplicationsStimuliresponsive2010}, thermal-responsive hydrogels~\cite{chenThermalresponsiveHydrogelSurface2010}, light-responsive molecules and microstructures~\cite{ichimuraLightDrivenMotionLiquids2000} or processes such as electrowetting~\cite{mugeleElectrowettingConvenientWay2005}.
\REV{Modelling the coupled problem of a thin film on a switchable substrate and as such the dynamics of the substarte as well as the film is far from being trivial but can be realized by a space and time dependent wettability pattern~\cite{grawitterSteeringDropletsSubstrates2021}}. 
While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{grawitterSteeringDropletsSubstrates2021,grawitterDropletsSubstratesOscillating2021,thieleGradientDynamicsModel2020} or limited to analysing the linear regime~\cite{sumanDynamicsThinLiquid2006}.

Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, starting from \REV{a slightly perturbed film to film} rupture \REV{and} long time morphology. 

@@

\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{oronLongscaleEvolutionThin1997,crasterDynamicsStabilityThin2009} 

@@

by means of a recently developed lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,zitzLatticeBoltzmannSimulations2021, zitzSwalbeJlLattice2022a}.
Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface. The mobility function 
$M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu}$
depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$
(for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ ($\gamma$ is the surface tension)
The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{bonnWettingSpreading2009, schwartzSimulationDropletMotion1998}. 

@@

where $f(\xi)=\xi^{-3} - \xi^{-2}$. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat}.

@@

directions~\cite{jalaliFabricationCharacterizationScalable2018,nagayamaIntermediateWettingState2020,dasSurfacedirectedSpinodalDecomposition2020}.

@@

On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads $(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{meckeThermalFluctuationsThin2005,zhangMolecularSimulationThin2019}. 

@@

\noindent {\it Results.} We first investigate how the rupture times depend on the parameters characterizing the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.

@@

The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. Hence we have 

>>>>

\noindent {\it Introduction}. Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{grossFluidFilmLubrication1980a,szeriFluidFilmLubrication2010,quereFluidCoatingFiber1999,dasilvasobrinhoStudyDefectsUltrathin1999,singhInkjetPrintingProcess2010,wijshoffDynamicsPiezoInkjet2010}. 
Moreover, the continuously growing interest for lab-on-a-chip devices~\cite{samieiReviewDigitalMicrofluidics2016,fockeLabonaFoilMicrofluidicsThin2010} as well as for printable electronics or printable photovoltaics~\cite{luechingerGraphenestabilizedCopperNanoparticles2008,ronsinFormationCrystallineBulk2022}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{buttAdaptiveWettingAdaptation2018}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{xinReversiblySwitchableWettability2010}, involving smart materials such as polymer brushes~\cite{stuartEmergingApplicationsStimuliresponsive2010}, thermal-responsive hydrogels~\cite{chenThermalresponsiveHydrogelSurface2010}, light-responsive molecules and microstructures~\cite{ichimuraLightDrivenMotionLiquids2000} or processes such as electrowetting~\cite{mugeleElectrowettingConvenientWay2005}.
\REV{Modelling the coupled problem of a thin film on a switchable substrate and as such the dynamics of the substrate as well as the film is far from being trivial but can be realized by a space and time dependent wettability pattern~\cite{grawitterSteeringDropletsSubstrates2021}}. 
While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{grawitterSteeringDropletsSubstrates2021,grawitterDropletsSubstratesOscillating2021,thieleGradientDynamicsModel2020} or limited to analysing the linear regime~\cite{sumanDynamicsThinLiquid2006}.

Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, starting from \REV{a slightly perturbed film to film} rupture \REV{and} long time morphology~\cite{konnurInstabilityMorphologyThin2000}. 

@@

\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009} 

@@

by means of a recently developed lattice Boltzmann (LB) scheme~\cite{zitzLatticeBoltzmannMethod2019,zitzLatticeBoltzmannSimulations2021, zitzSwalbeJlLattice2022a}.
Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface. 
The mobility function $M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu}$ depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$ (for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$)~\cite{huhHydrodynamicModelSteady1971, peschkaSignaturesSlipDewetting2019, fetzerQuantifyingHydrodynamicSlip2007, munchLubricationModelsSmall2005a}. 
Here, $\mu$ is the fluid dynamic viscosity.
The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ ($\gamma$ is the surface tension~\cite{wuHowChemicalPatterns2020}).
The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential~\cite{Deryaguin1940, DeryaguinChuraev1978}.
As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{bonnWettingSpreading2009, schwartzSimulationDropletMotion1998, youngIIIEssayCohesion1805, degennesWettingStaticsDynamics1985}. 

@@

where $f(\xi)=\xi^{-3} - \xi^{-2}$. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat, schwartzSimulationDropletMotion1998, mitlinDewettingSolidSurface1993, teletzkeHowLiquidsSpread1987}.

@@

directions~\cite{jalaliFabricationCharacterizationScalable2018, nagayamaIntermediateWettingState2020, dasSurfacedirectedSpinodalDecomposition2020}.

@@

On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads $(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{meckeThermalFluctuationsThin2005, zhangMolecularSimulationThin2019}. 

@@

\noindent {\it Results.} We first investigate how the rupture times depend on the parameters characterizing the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$~\cite{karguptaMorphologicalSelforganizationDewetting2002,karguptaInstabilityPatternFormation2000,nisatoExcitationSurfaceDeformation1999,karimPhaseSeparationUltrathin1998, succiLatticeBoltzmannEquation2001}.

@@

The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. 
Hence we have 
